---
title: R1. Model fitting 
subtitle: 2021 Skagit S/F Chinook Spawner Recruit Model fitting
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
---

***

This is version `r paste0('0.',format(Sys.time(), '%y.%m.%d'))`.

***

```{r set_options, echo = FALSE, message = FALSE}
options(width = 100)
knitr::opts_chunk$set(message = FALSE)
set.seed(123)
if(file.exists("cnt_time.txt")) {
  file.remove("cnt_time.txt")
}

mod_names <- data.frame(mod = rep(c("Beverton-Holt", "Beverton-Holt"), 1),
                        error = rep(c("MA1 & AR1", "AR1"), 1))
```

# Requirements
All analyses require the [R software](https://cran.r-project.org/) (v3.4.3) for data retrieval, data processing, and summarizing model results, and the [JAGS software](http://mcmc-jags.sourceforge.net/) (v4.2.0) for Markov chain Monte Carlo (MCMC) simulation. Please note that some of the R code below may not work with older versions of JAGS due to some changes in the ways that arrays are handled.

We also need a few packages that are not included with the base installation of R, so we begin by installing them (if necessary) and then loading them.

```{r load_pkgs, message = FALSE, warning = FALSE}
if(!require("here")) {
  install.packages("here")
  library("here")
}
if(!require("readr")) {
  install.packages("readr")
  library("readr")
}
if(!require("rjags")) {
  install.packages("rjags")
  library("rjags")
}
if(!require("loo")) {
  install.packages("loo")
  library("loo")
}
if(!require("ggplot2")) {
  install.packages("ggplot2")
  library("ggplot2")
}
if(!require("coda")) {
  install.packages("coda")
  library("coda")
}
if(!require("shinystan")) {
  install.packages("shinystan")
  library("shinystan")
}
if(!require("R2jags")) {
  install.packages("R2jags")
  library("R2jags")
}
if(!require("dclone")) {
  install.packages("dclone")
  library("dclone")
}
if(!require("snow")) {
  install.packages("snow")
  library("snow")
}
if(!require("rstan")) {
  install.packages("rstan")
  library("rstan")
}
if(!require("RColorBrewer")) {
  install.packages("RColorBrewer")
  library("RColorBrewer")
}
if(!require("gtools")) {
  install.packages("gtools")
  library("gtools")
}

management_unit <- "summer_fall"

## set directory locations
datadir <- here(paste(management_unit,"/","data",sep = ""))
jagsdir <- here(paste(management_unit,"/","jags",sep = ""))
analdir <- here(paste(management_unit,"/","analysis",sep = ""))
savedir <- here(paste(management_unit,"/","analysis/cache",sep = ""))
```

We also need a couple of helper functions.

```{r define_funcs}
## better round
Re2prec <- function(x, map = "round", prec = 1) {
  ## 'map' can be "round", "floor", or "ceiling"
  ## 'prec' is nearest value (eg, 0.1 means to nearest tenth; 1 gives normal behavior)
  if(prec<=0) { stop("\"prec\" cannot be less than or equal to 0") }
  do.call(map,list(x/prec))*prec
}

## wrapper function to fit JAGS models & rearrange output
fit_jags <- function(model, data, params, inits, ctrl, dir = jagsdir) {
  jm <- jags.model(file.path(jagsdir, model),
                   data,
                   inits,
                   ctrl$chains,
                   ctrl$burn,
                   quiet = TRUE)
  return(coda.samples(jm, params, ctrl$length, ctrl$thin))
}

#alternative wrapper to fit model in parallel; one chain per core
fit_jags2<-function(model,data,params,inits,ctrl,dir=jagsdir){
  cl <- makeCluster(3, type = "SOCK")
  inits2 <- jags.fit(data=data,
                     params=params, 
                     model=file.path(jagsdir, model), 
                     inits=inits,
                     n.chains=ctrl$chains,
                     n.adapt = 0,
                     n.update = 0,
                     n.iter = 0)$state(internal = TRUE)
  jm <- jags.parfit(cl=cl,
                    data = data,
                    params = params, 
                    model = file.path(jagsdir, model),
                    inits = inits2,
                    n.adapt = ctrl$burn*0.5,
                    n.update = ctrl$burn*0.5,
                    n.iter = ctrl$length-ctrl$burn,
                    thin = ctrl$thin,
                    n.chains = ctrl$chains
                    )
  stopCluster(cl)
  return(jm)
}

#generate summary stats file from MCMC object
sum_stats<-function(mcmclist){
  ESS<-apply(as.matrix(mcmclist),2,ess_bulk)
  Rhat<-apply(as.matrix(mcmclist),2,Rhat)
  summary_stats<-summary(mcmclist)
  summary_stats<-data.frame(summary_stats$statistics,summary_stats$quantiles,ESS,Rhat)
}

# functions for approximate LFO
# many functions modified from:
# https://github.com/paul-buerkner/LFO-CV-paper/blob/master/case-study-LFO-CV.Rmd

#load complete model fits & model refits with subset data
loadmodfits<-function(modelnames){
  mod_fits<-list(NULL)
  for(i in 1:length(modelnames)){
    mod_fits[[i]] <- readRDS(file.path(savedir,paste0(modelnames[i],"_y",n_forecasts+1,"_",run,".rds")))
    #mod_fits[[i]] <- readRDS(file.path(savedir,paste0("fit_",modelnames[i],".rds")))
  }
  return(mod_fits)
}

#refits
loadrefits<-function(refitname,N,L){
  numrefits<-N-L+1
  re_fits<-list()
  for(i in 1:numrefits){
     re_fits[[i]] <- readRDS(file.path(savedir,paste0(refitname,"_y",i,"_",run,".rds")))
  }
  return(re_fits)
}

# more stable than log(sum(exp(x))) 
log_sum_exp <- function(x) {
  max_x <- max(x)
  max_x + log(sum(exp(x - max_x)))
}

# more stable than log(mean(exp(x)))
log_mean_exp <- function(x) {
  log_sum_exp(x) - log(length(x))
}

# compute log of raw importance ratios
# sums over observations *not* over posterior samples
sum_log_ratios <- function(ll, ids = NULL) {
  if (!is.null(ids)) ll <- ll[, ids, drop = FALSE]
  - rowSums(ll)
}
# for printing comparisons later
rbind_print <- function(...) {
  round(rbind(...), digits = 2)
}

#function to extract log likelihood from fitted model
extract_log_lik<-function(m,esc_only,N,mod_fits){
  #extract pontwise log likelihoods
  tmp_lp <- as.matrix(mod_fits[[m]])
  ## extract pointwise likelihoods
  tmp_lp <- tmp_lp[,grepl("lp_", colnames(tmp_lp))]
  ## if numerical underflows, convert -Inf to 5% less than min(likelihood)
  if(any(is.infinite(tmp_lp))) {
    tmp_lp[is.infinite(tmp_lp)] <- NA
    tmp_min <- min(tmp_lp, na.rm = TRUE)
    tmp_lp[is.na(tmp_lp)] <- tmp_min * 1.05
  }
  if(esc_only =="Yes"){
    tmp_lp<-tmp_lp[,grepl("esc", colnames(tmp_lp))]
  }
  #get yrs assoc
  names_loglik<-data.frame(strsplit(colnames(tmp_lp),"\\[|\\]"))
  yrnames<-as.numeric(names_loglik[2,])
  
  loglik <- matrix(NA,ncol=N,nrow=dim(tmp_lp)[1])
  for(i in 1:N){
    if(!is.null(ncol(tmp_lp[,yrnames==i]))){
      loglik[,i] = apply(tmp_lp[,yrnames==i],1,sum)
    }else(loglik[,i] = tmp_lp[,yrnames==i])
  }
  return(loglik)
}

#function for printing out a read text file
processFile = function(filepath) {
  con = file(filepath, "r")
  while ( TRUE ) {
    line = readLines(con, n = 1)
    if ( length(line) == 0 ) {
      break
    }
    cat(paste0(noquote(line)),"\n")
  }
  close(con)
}

#=============================
#function to fit or load modelfits
#=================================
fit_load_mods<-function(models){
  
  models <- models
  
  ## empty list for fits
  mod_fits <- vector("list", n_mods*(n_forecasts+1))
  ## counter to index fitted jags models (33 in total: 3 models x 11 1 year ahead forecasts including upcoming
  ## return year)
  t <- 1
  for(n in 1:n_mods){
    ## counter to index data to feed model for year specific forecasts
    ## first forecast will be for 10 years prior to the most recent return year;
    ## last forecast will be current forecast for the upcoming return year
    c <- 0 
    model <- models[n]
    
    for(i in 1:(n_forecasts+1)){
      if(file.exists(file.path(savedir,paste(model,"_","y",i,"_",run,".rds",sep = "")))) {
        mod_fits[[t]] <- readRDS(file.path(savedir,paste(model,"_","y",i,"_",run,".rds",sep = "")))
        #summary_stats<-NULL
          #summary_stats<-sum_stats(mcmclist= mod_fits[[t]])
          #write.csv(summary_stats,file.path(savedir, paste(model,"_","y",i,"_summary_stats","_",run,".csv",sep = "")))
        c <- c + 1
        t <- t + 1
    } else { ## else, fit & save
          ## cnt & time stamp
          cat("Count =", t, "; Time =", round(((proc.time()-timer_start)/60)["elapsed"], 1), "\n",
              file="cnt_time.txt", append=TRUE)
          
          #range of years. Last year in range
          dat_yrs <- seq(yr_frst,(yr_last - n_forecasts + c),1)
          
          ## number of years of data
          n_yrs <- length(dat_yrs)
          
          
          ## get first & last years
          yr_frst_forecast <- min(dat_yrs)
          yr_last_forecast <- max(dat_yrs)
          
          ## get escapement data
          dat_esc_forecast <- dat_esc[which(dat_esc$year %in% dat_yrs),]
          
          ## log of escapement
          ln_dat_esc <- c(log(dat_esc_forecast$escapement),rep(NA,n_fore))
          
          
          ## get age data
          dat_age_forecast <- dat_age[which(dat_age$year %in% dat_yrs),]
          ## drop year col & first age_min+age_skip rows
          dat_age_forecast <- dat_age_forecast[-(1:(age_min+age_skip)),-1]
          
          ## add row(s) of NA's for forecast years
          if(n_fore > 0) {
            dat_age_forecast <- rbind(dat_age_forecast,
                                      matrix(0, n_fore, A,
                                             dimnames = list(n_yrs+seq(n_fore),colnames(dat_age_forecast))))
          }
          ## total num of age obs by cal yr
          dat_age_forecast[,"sum"] <- apply(dat_age_forecast, 1, sum)
          ## row indices for any years with no obs age comp
          idx_NA_yrs <- which(dat_age_forecast$sum<A, TRUE)
          ## replace 0's in yrs w/o any obs with NA's
          dat_age_forecast[idx_NA_yrs,(1:A)] <- NA
          ## change total in yrs w/o any obs from 0 to A to help dmulti()
          dat_age_forecast[idx_NA_yrs,"sum"] <- A
          ## convert class
          dat_age_forecast <- as.matrix(dat_age_forecast)
          
          ## get harvest data
          dat_harv_forecast <- dat_harv[which(dat_harv$year %in% dat_yrs),]
          ## drop year col & first age_max rows
          dat_harv_forecast <- c(dat_harv_forecast$catch,rep(NA,n_fore))
          
          ## get covariate data
          #dat_cvrs_forecast <- dat_cvrs[which(dat_cvrs$year <= yr_last + n_fore - age_min),1:4]
          ## drop year col
          #dat_cvrs_forecast <- dat_cvrs_forecast[,-1] 
          ## transform the covariates to z-scores
          #scl_cvrs_forecast <- scale(dat_cvrs_forecast) 
          ## total number of covariates
          #n_cov <- dim(dat_cvrs_forecast)[2] 
          
          ## ----jags_setup----------------------------------------------------------
          ## 1. Data to pass to JAGS
          dat_jags <- list(dat_age = dat_age_forecast,
                           ln_dat_esc = ln_dat_esc,
                           dat_harv = dat_harv_forecast,
                           A = A,
                           age_min = age_min,
                           age_max = age_max,
                           age_skip = age_skip,
                           n_yrs = n_yrs,
                           n_fore = n_fore)
          
          ## 2. Model params/states for JAGS to return
          ##    These are specific to the process model,
          ##    so we define them in 'par_jags' below.
          
            init_vals <- function() {
            list(alpha = 5,
                 #beta_inv = exp(mean(ln_dat_esc, na.rm = TRUE)),
                 beta=1/exp(mean(ln_dat_esc, na.rm=TRUE)),
                 pi_tau = 10,
                 pi_eta = rep(1,A),
                 pi_vec = matrix(c(0.020,0.219,0.581,0.179),
                                 n_yrs-age_min+n_fore, A,
                                 byrow = TRUE),
                 Rec_mu = log(1000),
                 Rec_sig = 0.1,
                 tot_ln_Rec = rep(log(1000), n_yrs - age_min + n_fore))
            }
            
            par_jags <- c("alpha","E_Rkr_a","mu_Rkr_a",
                          "beta",
                          "Sp","Rec","tot_ln_Rec","ln_RS","tot_Run",
                          "pi_eta","pi_tau","pi_vec",
                          "sigma_r","sigma_s","res_ln_Rec",
                          "lp_age","lp_esc","Run")

          
          #   init_vals <- function() {
          #   list(alpha = 5,
          #        beta_inv = exp(mean(ln_dat_esc, na.rm = TRUE)),
          #        pi_tau = 10,
          #        pi_eta = rep(1,A),
          #        pi_vec = matrix(c(0.020,0.219,0.581,0.179),
          #                        n_yrs-age_min+n_fore, A, 
          #                        byrow = TRUE),
          #        Rec_mu = log(1000),
          #        Rec_sig = 0.1,
          #        tot_ln_Rec = rep(log(1000), n_yrs - age_min + n_fore),
          #        phi_prior = 0.5,
          #        innov_1 = 0)
          #   }
          # 
          # 
          # if(model == "IPM_BH_AR"){
          #   ## params/states to return
          #   par_jags<- c("alpha","E_BH_a","mu_BH_a",
          #                 "beta",
          #                 "Sp","Rec","tot_ln_Rec","ln_RS",
          #                 "pi_eta","pi_tau","pi_vec",
          #                 "sigma_r","sigma_s","w","res_ln_Rec",
          #                 "lp_age","lp_esc","phi","Run"
          #                 )
          # 
          # }else{
          #   ## params/states to return
          #   par_jags <- c("alpha","E_Rkr_a","mu_Rkr_a",
          #                 "beta",
          #                 "Sp","Rec","tot_ln_Rec","ln_RS","tot_Run",
          #                 "pi_eta","pi_tau","pi_vec",
          #                 "sigma_r","sigma_s","res_ln_Rec","w","theta_res","phi",
          #                 "lp_age","lp_esc","Run"
          #                 )
          # 
          #   
          # }#endif
          # 
          
          
          ## set of multi-covariate models
          #cset <- colnames(scl_cvrs_forecast)
          #dat_jags$n_cov <- length(cset)
          #dat_jags$mod_cvrs <- scl_cvrs_forecast[1:(n_yrs-age_min+1), cset]
          
        
          ## fit model & save it
          # mod_fits[[t]] <- fit_jags(paste(model,".txt",sep = ""), dat_jags, par_jags,
          #                           init_vals_cov, mcmc_ctrl)
          mod_fits[[t]]<-fit_jags2(model=paste(model,".txt",sep = ""),
                             data=dat_jags,
                             params=par_jags,
                             inits=init_vals,
                             ctrl=mcmc_ctrl
                             )
          saveRDS(mod_fits[[t]], file.path(savedir,paste(model,"_","y",i,"_",run,".rds",sep = "")))
          summary_stats<-NULL
          summary_stats<-sum_stats(mcmclist= mod_fits[[t]])
          write.csv(summary_stats,file.path(savedir, paste(model,"_","y",i,"_summary_stats","_",run,".csv",sep = "")))
          c <- c + 1
          t <- t + 1
      }## end if
      
    }##next forecast year(i)
  }## next model(n)
  return(mod_fits)
}

```

# User inputs
We begin by supplying values for the following parameters, which we need for model fitting and evaluation.

```{r get_user_inputs}
## first & last years of fish data
yr_frst <- 1992
yr_last <- 2018



## min & max ad ult age classes
age_min <- 2
age_max <- 5

## years (if any) of age-comp to skip; see below
age_skip <- 0

## number of years ahead for run forecasts from the most recent year of data
n_fore <- 0

## number of recent year forecasts
n_forecasts <- 0

## first year of 1 step ahead forecast
#yr_begin <- 2011

## last year of 1 step ahead forecast
#yr_end <- 2020

## upper threshold for Gelman & Rubin's potential scale reduction factor (Rhat).
Rhat_thresh <- 1.1

## run sfck = summer/fall; spck = spring
run <- "sfck"

```

Next we specify the names of three necessary data files containing the following information:
 
 1. observed total number of adult spawners (escapement) by year;
 2. observed age composition of adult spawners by year;
 3. observed total harvest by year;

```{r get_filenames}
## 1. file with escapement data
## [n_yrs x 2] matrix of obs counts; 1st col is calendar yr
fn_esc <- paste("skagit","_",run,"_","esc.csv",sep = "")
paste("skagit","_",run,"_","esc.csv",sep = "")
## 2. file with age comp data
## [n_yrs x (1+A)]; 1st col is calendar yr
fn_age <- paste("skagit","_",run,"_","age.csv",sep = "")

## 3. file with harvest data
## [n_yrs x 2] matrix of obs catch; 1st col is calendar yr
fn_harv <- paste("skagit","_",run,"_","catch.csv",sep = "")
```

# Loading the fish data
Here we load in the first three data files and do some simple calculations and manipulations. First the spawner data:

```{r get_escapement_data}
## escapement
dat_esc <- read_csv(file.path(datadir, fn_esc))
## years of data
dat_yrs <- dat_esc$year

## number of years of data
n_yrs <- length(dat_yrs)

## log of escapement
ln_dat_esc <- c(log(dat_esc$escapement),rep(NA,n_fore))
```

Next the age composition data:

```{r get_age_data}
## age comp data
dat_age <- read_csv(file.path(datadir, fn_age))
## num of age classes
A <- age_max - age_min + 1

# ## drop year col & first age_min+age_skip rows
# dat_age <- dat_age[-(1:(age_min+age_skip)),-1]
#       
# ## add row(s) of NA's for forecast years
# if(n_fore > 0) {
#   dat_age <- rbind(dat_age,
#                    matrix(0, n_fore, A,
#                           dimnames =list(n_yrs+seq(n_fore),
#                                          colnames(dat_age))))
# }
# ## total num of age obs by cal yr
# dat_age[,"sum"] <- apply(dat_age, 1, sum)
# ## row indices for any years with no obs age comp
# idx_NA_yrs <- which(dat_age$sum<A, TRUE)
# ## replace 0's in yrs w/o any obs with NA's
# dat_age[idx_NA_yrs,(1:A)] <- NA
# ## change total in yrs w/o any obs from 0 to A to help dmulti()
# dat_age[idx_NA_yrs,"sum"] <- A
# ## convert class
# dat_age <- as.matrix(dat_age)
```

And then the harvest data:

```{r get_harvest}
## harvest
dat_harv <- read_csv(file.path(datadir, fn_harv))
## drop year col & first age_max rows
#dat_harv <- c(dat_harv$catch,rep(NA,n_fore))
```



# Specifying models in JAGS

Now we can specify the  model in JAGS. For this effort, we are only evaluating a Ricker model with gaussian process errors.

## Ricker with with gaussian process errors

```{r JAGS_RK_AR}
processFile(file.path(jagsdir, "IPM_RK.txt"))
```

***

# Fitting the models and generating the one year ahead forecasts
For the most recent 10 years (2011 - 2020), fit the model to data through year t-1 and generate a forecast for year t. 

Before fitting the model in JAGS, we need to specify the MCMC control parameters.

```{r jags_setup}
## 1. MCMC control params
mcmc_ctrl <- list(
  chains = 4,
  length = 5e5,#200000,
  burn = 2e5,#100000,
  thin = 400 #100
)
## total number of MCMC samples after burnin
mcmc_samp <- mcmc_ctrl$length*mcmc_ctrl$chains/mcmc_ctrl$thin
```

```{r start_timer, include = FALSE}
## start timer
timer_start <- proc.time() 
```


```{r jags_IO_base, warning = FALSE, cache = FALSE, message = FALSE}
## fit or load models
models=c("IPM_RK")
n_mods<-length(models)
mod_fits <- fit_load_mods(models=models)
```
